### 2015-04-28 thesis-proposal/problem.tex
update

Just as the Verse (i.e. ML) module system is organized around \emph{signatures} and \emph{modules}, the Verse metamodule system is organized around \emph{metasignatures} and \emph{metamodules}. Briefly summarized:
\begin{itemize}
\item A \emph{metasignature} specifies a set of type and operator constructors (tycons and opcons). Each is parameterized by a specified \emph{kind} of \emph{static value}. 
\item A \emph{metamodule}  defines static functions that govern the typechecking and  translation to the IL of each type and operator constructor specified in the  metasignature. 
\end{itemize}
\subsubsection{Example: Labeled Tuple Types via Metamodules}
\begin{figure}[hp!]
\begin{lstlisting}
metasignature LTUPLEX = metasig {
  ^\label{line:ltuplex-c-sig}^tycon c of LblTyOrdMap
  ^\label{line:intro-unlabeled-sig}^ana opcon intro_unlabeled (*of unit*)
  ^\label{line:intro-labeled-sig}^syn opcon intro_labeled of list(Lbl)
  ^\label{line:prj-sig}^syn opcon # of Lbl
  ^\label{line:conc-sig}^syn opcon + (*of unit*)
  ^\label{line:drop-sig}^syn opcon - of Lbl
}

syntax ltplx(L :: LTUPLEX, param :: LblTyOrdMap) at L.c(param) {
  fn ps => (* 
    - elaborates {e_1, ..., e_n} to L.intro_unlabeled e_1 ... e_n
    - elaborates {lbl1 => e_1, ..., lbln => e_n} to 
        L.intro_labeled [lbl1, ..., lbln] e_1 ... e_n
  *)
}

@@

  ana opcon intro_unlabeled {
  	(* translate intro operator to nested pairs *)
  	fn (ty :: Ty, param :: unit, args :: list(Arg)) => tycase(ty) {
        c(param) => (fold (zipExact (list_of_ltom param) args)
      	  `SCSS()ECSS`
          (fn (((lbl, ty), arg), cur_trans) => `SCSS(%{ECSSana(arg, ty)SCSS}, %ECSScur_transSCSS)ECSS`)
        )
      | _ => raise TyErr("...")
    }
  }

@@

Recall from Sec. \ref{sec:verse} that each valid external type has a translation, which is an internal type. The metamodule \lstinline{Ltuplex} defines a \emph{type translation schema}, a static function that determines the translations of types constructed by \lstinline{c} (each identified by the value of its parameter). Internal types are encoded as static values of type \lstinline{ITy}, which we assume has a TSL associated with it that allows us to use standard \emph{quasiquotation} syntax. Here, the  type translation schema on lines \ref{line:c-start} to \ref{line:c-end} generates nested internal products by folding over a list generated from the type parameter (in the standard way), returning the encoding of \lstinline{unit} for the empty case, and recursively nesting encodings of binary products otherwise using the standard \emph{unquote} operator, \lstinline{%}. 
For example, the encoding of the type translation of \lstinline{Paper} determined by this static function is:
\begin{lstlisting}[numbers=none]
`SCSStrans(ECSSrstring /SURL.+EURL/SCSS) * (trans(ECSSrstring /SURL([A-Z]+) (\d\d\d\d)EURL/SCSS) * unit)ECSS`
\end{lstlisting}
Notice that the translations of the component regular string types are not inserted directly, but are rather treated \emph{parametrically} using the \emph{translational form} \lstinline{SCSStrans(ECSSTSCSS)}. This will be key to the modularity principle that we plan to explore in the remaining work.

@@

\paragraph{Operator Constructors} On line \ref{line:intro-unlabeled-sig}, the metasignature \lstinline{LTUPLEX} specifies an operator constructor \lstinline{intro_unlabeled}. No parameter kind is listed, so the trivial parameter kind, \lstinline{unit}, is assumed, as indicated by the comment.

intro unlabeled is a trivially parameterized opcon, i.e. just one operator.

intro labeled is an opcon parameterized by a list of labels.

\# is an opcon parameterized by a single label.

ltpl is a TSM parameterized by a metamodule and a static value

LTuple uses ltpl as a TSL

\begin{lstlisting}
\end{lstlisting}

>>>>

Just as the Verse (i.e. ML) module system is organized around \emph{signatures} and \emph{modules}, the Verse metamodule system is organized around \emph{metasignatures} and \emph{metamodules}. Briefly summarized:
\begin{itemize}
\item A \emph{metasignature} specifies a set of type and operator constructors (tycons and opcons). Each is parameterized by a specified \emph{kind} of \emph{static value}. 
\item A \emph{metamodule}  defines static functions that govern the typechecking and  translation to the IL of each type and operator constructor specified in the  metasignature. 
\end{itemize}
\subsubsection{Example: Labeled Tuple Types via Metamodules}
\begin{figure}[hp!]
\begin{lstlisting}
metasignature LTUPLEX = metasig {
  ^\label{line:ltuplex-c-sig}^tycon c of LblTyOrdMap
  ^\label{line:intro-unlabeled-sig}^ana opcon intro_unlabeled of unit
  ^\label{line:intro-labeled-sig}^syn opcon intro_labeled of list(Lbl)
  ^\label{line:prj-sig}^syn opcon # of Lbl
  ^\label{line:conc-sig}^syn opcon + of unit
  ^\label{line:drop-sig}^syn opcon - of Lbl
}

syntax ltplx(L :: LTUPLEX, param :: LblTyOrdMap) at L.c(param) {
  fn ps => (* 
    - elaborates {e_1, ..., e_n} to L.intro_unlabeled(e_1; ...; e_n)
    - elaborates {lbl1 => e_1, ..., lbln => e_n} to 
        L.intro_labeled [lbl1, ..., lbln] e_1 ... e_n
  *)
}

@@

  ^\label{line:introul-start}^ana opcon intro_unlabeled {
  	(* translate intro operator to nested pairs *)
  	fn (ty :: Ty, param :: unit, args :: list(Arg)) => tycase(ty) {
        c(param) => (fold (zipExact (list_of_ltom param) args)
      	  `SCSS()ECSS`
          (fn (((lbl, ty), arg), cur_trans) => `SCSS(%{ECSSana(arg, ty)SCSS}, %ECSScur_transSCSS)ECSS`)
        )
      | _ => raise TyErr("...")
    }
  }^\label{line:introul-end}^

@@

Recall from Sec. \ref{sec:verse} that each valid external type has a translation, which is an internal type. The metamodule \lstinline{Ltuplex} defines a \emph{type translation schema}, a static function that determines the translations of types constructed by \lstinline{c} (each identified by the value of its parameter). Internal types are encoded as static values of type \lstinline{ITy}, which we assume has a TSL associated with it that allows us to use standard \emph{quasiquotation} syntax. Here, the  type translation schema on lines \ref{line:c-start} to \ref{line:c-end} generates nested internal products by folding over a list generated from the type parameter (in the standard way), returning the encoding of \lstinline{unit} for the empty case, and recursively nesting encodings of binary products otherwise using the standard \emph{unquote} operator, \lstinline{%}. 
For example, the encoding of the type translation of \lstinline{Paper} determined by this static function is:
\begin{lstlisting}[numbers=none]
`SCSStrans(ECSSrstring /SURL.+EURL/SCSS) * (trans(ECSSrstring /SURL([A-Z]+) (\d\d\d\d)EURL/SCSS) * unit)ECSS`
\end{lstlisting}
Notice that the translations of the component regular string types are not inserted directly, but are rather treated {parametrically} using the \emph{translational form} \lstinline{SCSStrans(ECSSTSCSS)}. This will be key to the modularity principle that we plan to explore in the remaining work.

@@

\paragraph{Operator Constructors} On line \ref{line:intro-unlabeled-sig} of Figure \ref{fig:ltuplex}, the metasignature \lstinline{LTUPLEX} specifies an operator constructor, \lstinline{intro_unlabeled} parameterized by static values of kind \lstinline{unit}. The qualifier \lstinline{ana} specifies that this opcon can only be used in analytic positions (i.e. where the expected type is known). For example, \lstinline{Ltuplex.intro_unlabeled} can be applied like this, because the return type of the function determines the expected type:
\begin{lstlisting}[numbers=none]
fun make_paper(conf : conf_t, title : title_t) : Paper => 
  Ltuplex.intro_unlabeled () (conf; title)
\end{lstlisting}
First, a static parameter value of the appropriate kind is supplied. After that, an \emph{argument list}, \lstinline{(conf; title)}, is supplied.

The metamodule \lstinline{Ltuplex} determines how this expression is typechecked and translated on lines \ref{line:introul-start}-\ref{line:introul-end}. Analytic opcons like \lstinline{intro_unlabeled} are defined with a static function that is given the type that the expression is being analyzed against, here \lstinline{Paper}, the operator parameter, here \lstinline{()}, and a list of \emph{argument interfaces}, which will allow it to ask the semantics to recursively typecheck and translate the provided arguments. The function must return an encoding of an  internal expression, which will become the translation of the external expression. Internal expressions are encoded as static values of kind \lstinline{IExp}, and again we assume a TSL has been defined so that we can use standard quasiquote syntax.
